My Publication Title --- Single Author

### 2025-07-07 paper/quarto_aaai/paper.tex
buff
De-emphasize training times as limitation of CT

\subsection{Our Proposed Objective}\label{our-proposed-objective}

Let \(\mathbf{x}_t^\prime\) for \(t=0,...,T\) denote a counterfactual
generated through gradient descent over \(T\) iterations as originally
proposed by \citet{wachter2017counterfactual}. In broad terms, searching
for CEs using gradient descent entails optimizing some form of an
objective that balances (1) the classification loss
\(\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})\), and (2) one
or more penalty terms \(\lambda_{i}\text{cost}_{i}(\cdot)\). The exact
specification of these penalties induces various properties in the
counterfactual outcomes, and tends to be the key feature that
distinguishes various gradient-based ``generators'' or ``explainers''
(CE algorithms) from the literature \citep{altmeyer2023explaining}.

@@

\emph{CT increases the training time of models.} CT can be more
time-consuming than conventional training regimes. Higher numbers of CEs
per iteration positively impact the quality of solutions, but they also
increase the amount of computations. In our experiments, grids with 270
settings took up to four hours for more demanding datasets on a
high-performance computing cluster with 34 2GB CPUs (see supplementary
appendix for details). Three factors attenuate this effect: (1) CT
inherently yields itself to parallel execution; (2) it amortizes the
cost of CEs for the training samples; (3) it retains its value when used
to fine-tune conventionally-trained models.

@@

>>>>

\subsection{Proposed Objective}\label{proposed-objective}

Let \(\mathbf{x}_t^\prime\) for \(t=0,...,T\) denote a counterfactual
generated through gradient descent over \(T\) iterations as originally
proposed by \citet{wachter2017counterfactual}. In broad terms, searching
for CEs using gradient descent entails optimizing some form of an
objective that balances (1) the classification loss
\(\text{yloss}(\mathbf{M}_\theta(\mathbf{x}),\mathbf{y})\), and (2) one
or more penalty terms \(\lambda_{i}\text{cost}_{i}(\cdot)\). The exact
specification of these penalties induces various properties in the
counterfactual outcomes, and tends to be the key feature that
distinguishes various gradient-based ``generators'' or ``explainers''
(CE algorithms) from the literature \citep{altmeyer2023explaining},
including all generators used in our experiments. We refer the reader to
the supplementary appendix for details.

@@

@@

\emph{CT increases the training time of models.} Just like adversarial
training, CT is more resource-intensive than conventional training
regimes. Higher numbers of CEs per iteration improve the quality of
learned representations, but they also increase the number of
computations. As our codebase is not performance optimized, grids of 270
settings took up to four hours for more demanding datasets in our
experiments, when ran on 34 2GB CPUs (see supplementary appendix). Other
than optimization, three factors mitigate this effect: (1) CT yields
itself to parallel execution; (2) it amortizes the cost of CEs for the
training samples; (3) it can be used to fine-tune conventionally-trained
models.
